---
title: "Visualize Data"
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
```

```{r}
mpg
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}

```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(mpg) + geom_point(aes(class, hwy))
```

## Your Turn 4

Make a histogram of the `hwy` variable from `mpg`.

```{r}

```

## Your Turn 5

Make a bar chart `hwy` colored by `class`.

```{r}

```

## Your Turn 6

Predict what this code will do. Then run it.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))
```

## Your Turn 7

The code from from the last exercise in the Console (the pane below this one). It will appear in the pane in the lower-right. Click "Export" and save the graph.

## Your Turn 8

The code below creates a scatterplot for the entire mpg dataset.
Modify the code to create a subplot for each value of drv.
Use ?facet_grid

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) 
```

## Your Turn 9

The code below creates a scatterplot for the entire mpg dataset.
Modify the code to create a subplot for each value of drv.
Use ?facet_grid

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) 
```

## Your turn 10

The code below creates a bar chart that shows the count of each class of vehicle in the dataset.

Make the chart interactive by passing the chart to ggplotly().

You will need to load the "plotly" library.
```{r}
chart = ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class))

chart
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```