---
title: "kable demo"
author: "Ari Lamstein"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

Experiment with rendering `head(mpg)` several ways.

1. As-is
2. As the output of `?knitr::kable`
3. As the output of `?knitr::kable` piped to `?kableExtra::kable_styling`

NOTE: Only load the kableExtra package while doing exercise #3. There seems to be a bug where when the kableExtra
package is loaded, it leads the regular kable package to return poorly styled output.

Then experiment with various options. In particular, with your partner, cycle 
through all options of the parameter `boostrap_options` in the `kable_styling` function.

If you have extra time, replace your called of `kable_styling` with another function in the `kableExtra` package
(such as `kable_minimal` or `kable_classic`). See this page for additional ideas: https://haozhu233.github.io/kableExtra/awesome_table_in_html.html

```{r}
# NOTE: Make sure you have all these packages installed!
library(knitr)
#library(kableExtra)

head(mpg)
```